\textbf{Too Beautiful to be Fake: Attractive Faces are Less Likely to be Judged as Artificially Generated}

### 2022-09-14 submission/manuscript.Rmd
Update manuscript.Rmd

<!-- Context -->
Literature shows that the context surrounding a stimulus often plays an important role in the assessment of its reality [a process henceforth referred to as *simulation monitoring*, @makowski2019phenomenal; @makowski2018cognitive]. The extensive search and processing of cues within ambiguous stimuli being and cognitively effortful [@michael2021source; @susmann2021persuasion].
Whereas individuals could rely on the content of the stimulus, making a reality judgement based solely on this information could be too cognitively effortful, especially if the stimulus is imbued with ambiguity [@michael2021source; @susmann2021persuasion]. As such, the Elaboration Likelihood Model (ELM) posits that people tend to draw on peripheral contextual cues, such as the source of the stimulus, credibility, authority and expertise of the source etc., to help facilitate their evaluation [petty1986elaboration]. Accordingly, such reliance on peripheral cues has been consistently reported across multiple studies, most recently with regards to people's trust in news articles reporting on the COVID-19 pandemic [@susmann2021persuasion;@michael2021source]. In the absence of these contextual information, however, what then  drives our beliefs of reality? (**Figure 1**).

@@

<!-- Internal features -->
Similar to contextual cues, the inter-individual characteristics of the assessor also impacts whether a stimulus is ultimately perceived to be real (or fake). Factors such as the individual's cognitive style, prior beliefs, as well as personality traits have been shown to significantly impact people's belief in ambiguous stimuli [@bryanov2021determinants; @ecker2022psychological; @sindermann2020short]. Specifically, individuals with higher levels of analytical reasoning have been found to be better at distinguishing real from fake stimuli, possibly since detection of false information requires careful scrutinization and deliberation [@pehlivanoglu2021role; @pennycook2019lazy]. Unsurprisingly, possessing prior knowledge or beliefs about the stimulus at hand also influences one's perception of it; biasing us to attend only to information in line with our predispositions (what is known as **confirmation bias**) [@britt2019reasoned]. For instance, in a study conducted by [@tucciarelli2020], participants who were told about the presence of social agents reported lower levels of trust relative to those who were not relayed any information. Furthermore, dispositional traits, such as high levels of narcissism and low levels of openness and conscientiousness have been associated with greater susceptibility to fake news [@piksa2022cognitive; @sindermann2020short].

<!-- Reactive features -->
Aside from stimulus- and individual-related characteristics, it is possible that simulation monitoring is driven by an interaction between the two, i.e., by the reaction associated with the experience of a given stimulus. One way in which this is exemplified is via emotions. For instance, in the field of virtual reality (VR), emotional features have been shown to increase the virtual domain's presence, i.e., the extent to which the the system is able to deliver a convincing (real-like) environment with which an agent can interact [@sanchez2005presence; @makowski2017being]. This hence implies that the inclusion of emotional features could increase people's belief in simulations. Indeed, several studies have suggested that heightened stimulus emotionality is associated with greater belief in fake, but not real, information [@martel2020reliance]. On the contrary, research relating emotional works of fiction to simulation monitoring have instead found emotionality to be associated with reduced perceptions of realness [@makowski2019phenomenal], i.e., emotional stimuli were more likely perceived to be fake as opposed to real. However, since participants were told the descriptions corresponding to some of the stimuli were fake, the latter finding could be understood as a result of the participants engaging in spontaneous fictional appraisal, an emotion regulation mechanism in which individuals lower their perception of realness of the given stimulus by endowing it with a fictional context [@sperduti2017distinctive; @sperduti2016paradox]. Moreover, when the stimuli was self-relevant, studies show that the experienced intensity of emotions was enhanced [@sperduti2016paradox; @goldstein2009pleasure]. This hence implies that self-relevance could modulate the relationship between the emotionality of a stimulus and our perceptions of it's realness. Additionally, other features that drive individuals' evaluation of a stimulus, such as familiarity, have also been shown to correlate positively with the perception of it as truth [@begg1992dissociation]

>>>>

<!-- Context -->
Literature shows that the context surrounding a stimulus often plays an important role in the assessment of its reality [a process henceforth referred to as *simulation monitoring*, @makowski2019phenomenal; @makowski2018cognitive]. The extensive search and processing of cues within ambiguous stimuli being an increasingly complex and cognitively effortful strategy [@michael2021source; @susmann2021persuasion], people tend to draw on peripheral contextual cues (**Figure 1**), such as the source of the stimulus, and its credibility, authority and expertise, to help facilitate their evaluation [petty1986elaboration; @susmann2021persuasion; @michael2021source]. However, the atomization and decontextualization of information allowed by online social media (where text snippets or video excerpts are mass-shared with little context) can render this task difficult. In the absence of contextual information, what drives our beliefs of reality?

@@

<!-- Internal features -->
The evidence shows that inter-individual characteristics play a role in the formation of beliefs of reality, with factors such as cognitive style, prior beliefs, and personality traits significantly impacting simulation monitoring [@bryanov2021determinants; @ecker2022psychological; @sindermann2020short]. For instance, individuals with higher levels of analytical reasoning have been found to better discriminate real from fake stimuli [@pehlivanoglu2021role; @pennycook2019lazy]. Prior knowledge or beliefs about the stimulus influences one's perception of it by biasing the attention deployment towards information in line with our expectations [@britt2019reasoned], and dispositional traits, such as high levels of narcissism and low levels of openness and conscientiousness, have been associated with greater susceptibility to fake news [@piksa2022cognitive; @sindermann2020short].

<!-- Reactive features -->
Beyond stimulus- and individual-related characteristics, evidence suggests that the interaction between the two, i.e., the subjective reaction associated with the experience of a given stimulus, contributes to simulation monitoring decisions. For instance, the intensity of experienced emotions have been shown to increase the sense of presence, i.e., the extent to which one feels like "being there", as if the movie or the VR environment was real [@sanchez2005presence; @makowski2017being], and beliefs that the stimuli were fake led to emotion down-regulation [@makowski2019phenomenal; sperduti2017distinctive]. However, @martel2020reliance reported that heightened stimulus emotionality was associated with greater belief in fake, but not real, information. Other factors, such as the stimulus' perceived self-relevance [@sperduti2016paradox; @goldstein2009pleasure], as well as familiarity  [@begg1992dissociation], could also play a role in our processing and reaction to real vs. non-real material.
